---
title: "Untitled"
author: "Linus Hof"
date: "2023-04-27"
output: md_document
---

# Resources

The following list of resources may help you with the tasks: 

- [Slides Session 2](https://www.moodle.tum.de/pluginfile.php/4527976/mod_resource/content/2/session_2_software.pdf)
- [Software Primer](https://github.com/linushof/BayesIntro/blob/main/setup/software_primer.md)

# Exercise 1: Creating and cloning the course repository

1. Create a new *GitHub* repo which you will retain throughout this course.
2. Clone the repo to your machine by creating a new *RStudio* project with version control.
3. In your local repo, create a new folder `exercises` and a new R script `session_2_exercise.R`.
4. Push the `exercise` folder and `session_2_exercise.R` to *GitHub*.
5. Write the code for the remaining exercises in `session_2_exercise.R` and commit and push regularly. 
Use commented out code (`#`) to add notes to your script and document your code.

# Exercise 2: Working with vectors and functions

## 2.1

Consider the following lines of code:  

```{r}
x <- runif(100, 1, 100)
mean_x <- round(mean(x), 2)
```
1. Try to understand what the code is doing by running it and investigating the outputs.
2. Check your intuition by looking up the help pages of the functions `runif`, `mean`, and `round`.
3. Run the above code repeatedly and check `mean_x` after each run. Why does the value change?
4. Search the internet for a command that allows you to obtain the same value for `mean_x` every time you rerun the two code lines. 
Fix the code to obtain the desired results. 


## 2.2 

1. Generate a vector `x` that entails all integers between 1 and 1000.
2. Find at least one more way to generate the same vector `x`.
3. Generate a vector `y` such that the output of `x + y` returns a vector that only contains 1000s. 
4. Assume `x` and `y` contain the amounts in dollars that 2000 different people possess. 
How much money do all people possess together?

## 2.3 

The following code draws 10,000 random values from a beta distribution with the shape parameters `a = 4` and `b = 4` and multiplies them by a scale parameter `scale = 15,000`. 

```{r}
n <- 1e4
scale <- 1.5e4
income <- round( rbeta(n=n, shape1=2, shape2=12) * scale, 2)
```

Let's assume these simulated values represent the gross income in US$ of 10,000 people. 
The following lines of code plot a histogram, which displays the counts of people that obtain an income in the respective range. 

```{r}
library(ggplot2) # only load (run) once
```

```{r}
# Plot the resulting curve
ggplot(data.frame(x = income), aes(x=x)) +
  geom_histogram(color = "#0065BD", fill = "#0065BD", alpha=0.5, bins = 100) +
  scale_x_continuous(breaks = seq(0, scale, 1e3)) + 
  labs(x = "Gross income", 
       y = "Counts") + 
  theme_minimal()
```

1. From what you know or assume about how income is distributed in a population, do you think the simulated data is realistic?
Change the scale and shape parameters in the function `rbeta` to simulate new data and generate a new plot. 
Play a bit around and see what the parameters do to your data by investigating the resulting histograms.
Try to use the above code to simulate the data which reflects your knowledge or assumptions about income distributions.
(*Tipp: If you overwrite the object `income`, the code for the plot must not be changed. If you create a new object, substitute `income` in the function `ggplot()` with the new object.*)

2. Use the income distribution `income` that you came up with to compute each person's share of the total income of all people. 

3. Investigating the individual share values may not be very enlightening. 
You may instead create more comprehensive summaries of income distributions as does the following code. 
By investigating and running the code, try to understand it and comment the code lines and/or blocks using commented out lines (`#`).  

```{r}
income_s <- sort(income)
group <- c("Lower 1%", "Lower 50%", "Top 10%", "Top 1%")
p <- c(.1, .5, .9, .99)

boundary <- round(income_s[p*n], 0)

low10_m <- mean( income_s[c(1:(.1*n))] )
low50_m <- mean( income_s[c(1:(.5*n))] )
top10_m <- mean( income_s[c((.9*n):n)] )
top1_m <- mean( income_s[c((.99*n):n)] )

means <-  round( c(low10_m, low50_m, top10_m, top1_m) , 0)

income_summary <- data.frame(group, boundary, means)
income_summary
```

4. Considering the above summary table and what you know or assume about how income is distributed in a population, do you still think the simulated data is realistic?
Change the scale and shape parameters in the function `rbeta` to simulate new data and generate new plots and summaries. 

# Exercise 3: Working with data frames 

1. Create a data frame with 5 variables and 5 rows. 
2. Computes the arithmetic mean of each variable in the data set and store all 5 means in one vector.
3. *Bonus*: If you copied and pasted the code for each variable, try to come up with a solution, that omits copy and pasting and does the entire operation in one line.
Make sure that the output shows only the computed names and not also the respective variable names.

Below, you can see the first 10 rows of the `diamonds` data set from the *tidyverse*. 

```{r}
head(diamonds, 10)
```

3. Load the tidyverse package with `library(tidyverse)` to obtain the full data set.
Get an overview of the data set with the commands `names(diamonds)` and `View(diamonds)`.

4. Try to obtain all exemplars (rows) that have an ideal cut. First try to use the matrix notation together which the function `which`. 
Then visit the help pages for the function `filter` from the *dplyr* package that is attached to the *tidyverse*. 
Use the filter function to solve the problem and also play a bit around with *dplyr*'s function `select`.




